---
title: "Network Analytics PreModule"
author: "Zachary Long"
date: "2023-10-27"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 3 Network structure visualization

Install the igraph package for R (suggested) or Python. To do so, in R type:1
1 # Download and install the package
2 install . packages (" igraph ")
1. Plot the network using the information in the  le graph subset rank1000.txt. Note that
this is not the complete network, but only a subset of edges between top-ranked products.
By visualizing the graph, you get an idea of the structure of the network you will be working
on. In addition to plotting, comment on anything interesting you observe.
Hints:
• Refer to https://kateto.net/netscix2016.html for a tutorial on igraph and R basics.
• You may  nd it useful to treat this data  le as being in ncol format in igraph.
• It may be simplest to treat the network as undirected for the purposes of visualization
(since directed arrows can add a lot of visual clutter in a graph of this size).
• Playing with the size, color, and layout of objects may make the network easier to
visualize. When plotting you can start with layout=layout.auto and then experiment
with other options. layout=layout.kamada.kawai generally gives good results.
2. Now, use the  le graph subset rank1000 cc.txt to plot only the largest connected compo-
nent in the above network. You should be able to reuse your code from above on the new
data.
```{r }
install.packages("igraph", repos = "http://cran.us.r-project.org")


library(igraph)
library(ggplot2)

setwd("/Users/zacharylong/Library/CloudStorage/Dropbox/NYU MSBA/Module 4/Network Analytics/Pre-Module")

```


```{r, fig.width=8, fig.height=8 }
# Let's take our text file of space-separated edges and put it into an iGraph graph
graph_subset_rank1000 <- read.graph(file="data/graph_subset_rank1000.txt", format ="ncol", directed = FALSE)
graph_subset_rank1000

# How many nodes?
vcount(graph_subset_rank1000)

# How many edges?
ecount(graph_subset_rank1000)

# Get nodes
V(graph_subset_rank1000)

# Get edges
E(graph_subset_rank1000)

# Is the graph directed?
is.directed(graph_subset_rank1000)

# Plot using the tutorial starter code
plot.igraph(graph_subset_rank1000)

```

Plot with the layout.auto option from the directions

```{r, fig.width=8, fig.height=8 }

plot(graph_subset_rank1000, layout = layout.auto(graph_subset_rank1000))

```
Looking at the shape, we can see an internal cluster of nodes and a ring of outer nodes. Plot a histogram of the distribution of the degrees to examine the structure.

```{r}
degree_dist <- degree(graph_subset_rank1000)
hist(degree_dist)

```

Based on the visual structure, examine the clustering coefficient, or the measure of how the nodes cluster together.

```{r}
transitivity(graph_subset_rank1000, type="global")
```

Examine the avergae path length, to see how many hops it takes to get from one node to another on average. This number is slightly larger than our "6-degrees of separation" from the earlier assignment

```{r}
avg_path_length <- average.path.length(graph_subset_rank1000)
print(avg_path_length)
```

From the tutorial code, check the betweenness centrality for all nodes, which is the extent a node acts as a bridge between other nodes in the network.

```{r}
btw <- betweenness(graph_subset_rank1000)

# I will round it to 2 digits just so that it's easier to read
btw <- round(btw, 2)
# Just show the head() for the knit output file.
head(btw)
```

```{r}
# Sort the betweenness scores
sort(btw)
```

```{r}
# Get the unique values of betweenness scores, and their count; we'll use a dataframe so the result is easier to read
data.frame(table(btw))
```

```{r}
# For illustrating how to use ggplot, let's just make a nonsensical scatterplot of betweenness against itself
ggplot(data.frame(btw), aes(x=btw, y=btw)) + geom_point()
```

```{r, fig.width=8, fig.height=8}
# Refine the plot

# Calculate the degree of each node
node_degree <- degree(graph_subset_rank1000)

# Create a data frame
df <- data.frame(Degree = node_degree, Betweenness = btw)

# Establish a threshold for the plot
threshold <- quantile(btw, 0.95, na.rm = TRUE)

# Create ggplot
ggplot(df, aes(x=Degree, y=Betweenness)) +
  geom_point(aes(color = Degree)) +  # Color by Degree
  geom_smooth(method = 'loess') +    # Add smoothed line
  geom_text(aes(label=ifelse(Betweenness > threshold, as.character(rownames(df)), '')), hjust=1.5, vjust=0.5) +  # Label extreme points
  labs(title = "Scatterplot of Betweenness Centrality vs. Node Degree",
       x = "Node Degree",
       y = "Betweenness Centrality")

```

Use kamada kawai layout from the assignment suggestions
```{r, fig.width=8, fig.height=8}
plot(graph_subset_rank1000, layout = layout.kamada.kawai(graph_subset_rank1000), vertex.size = 5, vertex.color = "blue", edge.color = "grey")
```

Extract and plot the largest connected component

```{r, fig.width=8, fig.height=8}
# Find the connected components in the graph
comps <- components(graph_subset_rank1000)

# Which is the largest component
print("Largest connected component: ")
which.max(comps$csize)

# Get the betweenness scores for node 57
btw[c("57")]

# Find the indices of the nodes that belong to the largest component
print("Indices of the nodes that belong to the largest connected component: ")
which(comps$membership == which.max(comps$csize))

# Create a graph that includes only the nodes that belong to the largest connected component
largest_comp <- induced_subgraph(graph_subset_rank1000, which(comps$membership == which.max(comps$csize)))

# Add colors and plot the graph
print("Graph of the largest connected component: ")
plot(largest_comp, layout = layout.kamada.kawai(largest_comp), vertex.size = 5, vertex.color = "red", edge.color = "grey")

```
Plot again

```{r, fig.width=8, fig.height=8}
# Create a data frame for edges
edges_df <- get.data.frame(largest_comp, what="edges")

# Create a data frame for vertices
vertices_df <- data.frame(id = V(largest_comp)$name)

# Calculate layout
layout_matrix <- as.data.frame(layout.kamada.kawai(largest_comp))
colnames(layout_matrix) <- c("x", "y")

# Add layout coordinates to vertices data frame
vertices_df <- cbind(vertices_df, layout_matrix)

# Map layout coordinates to edges data frame
edges_df <- merge(edges_df, vertices_df, by.x="from", by.y="id")
edges_df <- merge(edges_df, vertices_df, by.x="to", by.y="id", suffixes = c("_from", "_to"))

# Create ggplot
ggplot() +
  geom_segment(data=edges_df, aes(x=x_from, y=y_from, xend=x_to, yend=y_to), color="grey50") +
  geom_point(data=vertices_df, aes(x=x, y=y), color="red", size=3) +
  geom_text(data=vertices_df, aes(x=x, y=y, label=id), vjust=1.5, hjust=0.5) +
  labs(title="Largest Connected Component") +
  theme_void()

```

#4 Data Analysis

For the rest of the assignment, use the complete graph contained in the file graph complete.txt
and the title file id to titles.csv. It will be in in your best interest to using a programming
language such as R or Python.
If you face computational challenges analyzing the larger data set graph complete.txt, you may
contact your TA for permission to use the data set graph subset rank1000.txt instead, with a
brief explanation about what barriers you faced using the big data option.
Note: Here, we are working with a directed graph. For example, the \Grapes of Wrath" prod-
uct page might highlight a co-purchase link to \East of Eden", but the \East of Eden" product
page might not necessarily link back to the \Grapes of Wrath" product page, and might instead
link to \The Winter of Our Discontent". Each product can have multiple inbound or outbound
edges.

# 4.1. Plot the out-degree distribution of our dataset (x-axis number of similar products, y-axisnumber of nodes). 
That is, for each product a, count the number of outgoing links to another
product page b such that a > b.
Hint: The following steps will outline one way to approach this problem.
(a) Start by calculating the out-degree for each product. You may use the table command
in R or a dict in Python to compute the number of outbound links for each product.
(b) You can then apply the same process you just used so that you can count the number of
products (nodes) that have a particular number of outgoing links. This is the out-degree
distribution.
(c) Once you are done, you can use the default plotting environment in R, ggplot2 in R,
or matplotlib3 in Python to plot the distribution. Note that you can avoid step (b) if
you use the geom density() function in ggplot or the hist() method in matplotlib.
However, you may approach this any way you wish.

```{r}
# Calculate the out-degree for each product

library(ggplot2)
library(dplyr)

# Read in the complete graph of edges
edges <- read.table("data/graph_complete.txt", header = FALSE, sep = " ")

# The edges of the graph in the form from -> to. Each line is an edge, with the origin node and destination node separated by a space.
colnames(edges) <- c("From", "To")

# Calculate the out-degree using the table command in the instructions
out_degree <- table(edges$From)
out_degree_distribution <- table(out_degree)

# View the distribution of Out-Degrees from table()
print("Distirbution of Out-Degrees from table():")
out_degree_distribution

# Use iGraph package's graph_from_data_frame() to check Out-Degrees using directed=TRUE
igraphDegrees <- graph_from_data_frame(edges, directed = TRUE)
out_degrees_igraph <- degree(igraphDegrees, mode = "out")

# Convert the iGraph results to table
out_degrees_igraph_table <- table(out_degrees_igraph)
out_degrees_igraph_table

# View the out-degrees from iGraph
print("iGraph function shows the product node and the number of out-degrees from it:")
head(out_degrees_igraph)

# Add the zero values back to the table() output
# Identify All Unique Nodes
all_nodes <- unique(c(edges$From, edges$To))

# Find Nodes with Zero Out-Degrees
zero_out_degree_nodes <- setdiff(all_nodes, unique(edges$From))

# Add Zero Out-Degrees to Table
zero_out_degree_table <- table(rep(0, length(zero_out_degree_nodes)))
out_degree_distribution_adjusted <- table(c(as.integer(out_degree), rep(0, length(zero_out_degree_nodes))))

# View the adjusted distribution of Out-Degrees
print("Adjusted Distribution of Out-Degrees from table():")
out_degree_distribution_adjusted

```
Commentary on output: table() allows us to create a table for plotting the distribution of out-degrees, which is summarized here:
[1] "Disitrbution of Out-Degrees from table():"
out_degree
    1     2     3     4     5 
38400 55689 79244 97275 90958 

iGraph also has a function for this:
out_degrees_igraph
    0     1     2     3     4     5 
 5421 38400 55689 79244 97275 90958 

Comparing the two outputs from table() to using the degree() function from igraph which examines the entire network, we can see that table() did not account for the zero out-degree products. We can calculate the total number of nodes/products and add that back into the output.

Adjusted Distribution of Out-Degrees from table():

    0     1     2     3     4     5 
 5421 38400 55689 79244 97275 90958 

Using the igraph package's graph_from_data_frame() function we can see the number of out-degrees for each product node:
[1] "iGraph function shows the product node and the number of out-degrees from it:"
140890 204319   1046 170475 183024 231749 
     4      4      2      1      3      2 

Next for 4.1, Plot the Out-degree Distribution from table()
```{r }
# Convert the table to dataframe for ggplot
out_degree_distribution_df <- data.frame("OutDegree" = as.numeric(names(out_degree_distribution_adjusted)), 
                                         "Count" = as.numeric(out_degree_distribution_adjusted))

# Plot the distribution of out-degrees using ggplot
ggplot(out_degree_distribution_df, aes(x=OutDegree, y=Count)) +
  geom_point() +
  geom_line(group=1) +
  ggtitle("Out-degree Distribution") +
  xlab("Number of Outgoing Links") +
  ylab("Number of Nodes")



```
GGPlot Geom_Density plot for Out-Degrees

```{r}
# Re-convert the table to a dataframe for geom_density
out_degree_df <- data.frame("OutDegree" = as.numeric(out_degree))

# Revise this with the new zero degree nodes
# # Plot using ggplot2 with geom_density
# ggplot(out_degree_df, aes(x=OutDegree)) +
#   geom_density(fill="blue", alpha=0.5) +
#   ggtitle("Out-degree Distribution with geom_density()") +
#   xlab("Number of Outgoing Links") +
#   ylab("Density")

zero_out_degree_df <- data.frame("OutDegree" = rep(0, length(zero_out_degree_nodes)))

out_degree_df_adjusted <- rbind(out_degree_df, zero_out_degree_df)


# Step 1: Create a Dataframe for Zero Out-Degrees
zero_out_degree_df <- data.frame("OutDegree" = rep(0, length(zero_out_degree_nodes)))

# Step 2: Combine the Two Dataframes
out_degree_df_adjusted <- rbind(out_degree_df, zero_out_degree_df)

# Step 3: Plot with ggplot2
ggplot(out_degree_df_adjusted, aes(x=OutDegree)) +
  geom_density(fill="blue", alpha=0.5) +
  ggtitle("Adjusted Out-degree Distribution with geom_density()") +
  xlab("Number of Outgoing Links") +
  ylab("Density")

```

Plot with igraph dataframe
```{r}
# plot of out degree distribution
df_out_degree <- data.frame(x = out_degrees_igraph_table)

ggplot(data = df_out_degree, aes(x=as.numeric(x.out_degrees_igraph))) +
  geom_density(color = 'blue') +
  labs(
    title = "Density Plot of Out-Degree Distribution with iGraph Data",
    x = "Number of Out-Degree",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0), name = "Degrees", breaks = 0:5, limits = c(0, 5))+
  theme_bw()


```


#4.2 In-Degree Distribution Plot

```{r}
# Read the table for the "To" direction.
in_degree <- table(edges$To)

# Calculate in-degree distribution with table()
in_degree_distribution <- table(in_degree)

# As earlier, account for zero in-degrees with table()
# Find the zero in-degree nodes
zero_in_degree_nodes <- setdiff(all_nodes, unique(edges$To))

# Add to our table
in_degree_adjusted <- table(c(as.integer(in_degree), rep(0, length(zero_in_degree_nodes))))

# Update dataframe for ggplot
in_degree_distribution_adjusted_df <- data.frame("InDegree" = as.numeric(names(in_degree_adjusted)), 
                                                 "Count" = as.numeric(in_degree_adjusted))

# Step 5: Plot with ggplot2
ggplot(in_degree_distribution_adjusted_df, aes(x=InDegree, y=Count)) +
  geom_point() +
  geom_line(group=1) +
  ggtitle("In-degree Distribution with table()") +
  xlab("Number of Incoming Links") +
  ylab("Number of Nodes")


```
Geom_Density for In-Degrees

```{r}
# Create a new data frame for geom_density()
in_degree_for_density <- rep(in_degree_distribution_adjusted_df$InDegree, 
                             in_degree_distribution_adjusted_df$Count)

in_degree_for_density_df <- data.frame("InDegree" = in_degree_for_density)

# Generate the density plot using ggplot2
ggplot(in_degree_for_density_df, aes(x=InDegree)) +
  geom_density(fill="blue", alpha=0.5) +
  ggtitle("Adjusted In-degree Distribution with geom_density()") +
  xlab("Number of Incoming Links") +
  ylab("Density")

```
Is the distribution different? Comment on what you observe.
As we saw earlier in 4.1 for the out-degree, each product has a maximum of five outbound links with fewer having a single or two links.
The in-degree plot shows that their are a large number of products with very few inbound links to other products.

```{r}
# As before with out-dgerees, use iGraph version of in-degree using degree() to check in degrees for the entire graph.

# Calculate in-degrees using igraph's degree() mode in
in_degrees_igraph <- degree(igraphDegrees, mode = "in")

# Convert the iGraph results to table
in_degrees_igraph_table <- table(in_degrees_igraph)

# Convert to dataframe to plot of out degree distribution
in_degree_igraph_df <- data.frame(x = in_degrees_igraph_table)

# Convert data type from factor to numeric
in_degree_igraph_df <- in_degree_igraph_df %>%
  mutate(x.in_degrees_igraph = as.numeric(levels(x.in_degrees_igraph))[x.in_degrees_igraph])

# Plot with ggplot geom_density()
ggplot(data = in_degree_igraph_df, aes(x=x.in_degrees_igraph)) +
  geom_density(aes(fill = "Desnsity"), alpha = 0.5, color = 'red') +
  labs(
    title = "Density Plot of In-Degree Distribution with igraph",
    x = "Number of In-Degree",
    y = "Density"
  ) +
   scale_x_continuous(
    breaks = seq(min(in_degree_igraph_df$x.in_degrees_igraph),
                max(in_degree_igraph_df$x.in_degrees_igraph), by = 50)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  guides(fill="none")


```


#4.3 Transform and convert to log

```{r}
# Replace 0 with 0.1 to make it compatible with log
in_degree_distribution_adjusted_df$InDegree[in_degree_distribution_adjusted_df$InDegree == 0] <- 0.1

```

```{r}
# Plot using ggplot2 with log-scaled x-axis
ggplot(in_degree_distribution_adjusted_df, aes(x=InDegree, y=Count)) +
  geom_point() +
  geom_line(group=1) +
  ggtitle("In-degree Distribution with Log-Scaled x-axis from table()") +
  scale_x_log10() +
  xlab("Number of Incoming Links (Log Scale)") +
  ylab("Number of Nodes")

```

```{r}
# Do the same log transformation to the igraph version of the calculations
# Replace 0 with 0.1 to make it compatible with log
in_degree_igraph_df$x.in_degrees_igraph[in_degree_igraph_df$x.in_degrees_igraph == 0] <- 0.1

# Plot using ggplot2 with log-scaled x-axis for igraph data

ggplot(data = in_degree_igraph_df, aes(x=as.numeric(x.in_degrees_igraph))) +
  geom_density(aes(fill = "Density"), alpha = 0.5, color = 'red') +
  labs(
    title = "Enhanced Density Plot of In-Degree (log) Distribution",
    x = "Number of In-Degree in Log Scale",
    y = "Density"
  ) +
  scale_x_log10() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  guides(fill="none")



```


4.3 Comment on what you observe.

By transforming the x-axis to a log scale, it is easier to see the distribution of in-degrees with the previous sharp log shape of the graph having the data points bunched together near zero.

#4.4 Compute the average number of in-bound co-purchase links, the standard deviation, and the maximum
```{r}
# Calculate the average number of inbound links using the calculation from table()
average_in_degree_table <- mean(as.numeric(in_degree))

# Calculate the standard deviation of inbound links using the calculation from table()
sd_in_degree_table <- sd(as.numeric(in_degree))

# Find the maximum number of inbound links using the calculation from table()
max_in_degree_table <- max(as.numeric(in_degree))

# Print the table() results
cat("Average number of inbound links using table():", average_in_degree_table, "\n")
cat("Standard deviation of inbound links using table():", sd_in_degree_table, "\n")
cat("Maximum number of inbound links using table():", max_in_degree_table, "\n")

# Also use the igraph results for mean, std dev, and max
cat("\n", "Compare table() and igraph results", "\n", "\n")
# Calculate the average number of inbound links using igraph
average_in_degree <- mean(in_degrees_igraph)

# Calculate the standard deviation of inbound links using igraph
sd_in_degree <- sd(in_degrees_igraph)

# Find the maximum number of inbound links using igraph
max_in_degree <- max(in_degrees_igraph)

# Print the results
cat("Average number of inbound links using igraph:", average_in_degree, "\n")
cat("Standard deviation of inbound links using igraph:", sd_in_degree, "\n")
cat("Maximum number of inbound links using igraph:", max_in_degree, "\n")


```
4.4 Comment on the result.

I have included the output of both the earlier table() and igraph calculation for a different prespective on the results. igraph takes into account the entire graph, including zeroes, and while we corrected for that earlier, this output shows two different stories. The igraph results show a slightly lower average number of inbound links due to the inclusion of the zeros. Both align with our initial exploratory graphs showcasing the maximum out degrees being 5 and the logarithmic nature of in degrees. The table() results are slightly higher due to not accounting for zero nodes, but excluding the isolated nodes may be worth considering from a operating perspective since they are isolated from discovery on the Amazon website. Using these statistics, such as maximum, can find the most popular products that have the most links pointing to them.

#4.5 Report the names of the 10 products with the most inbound co-purchase links

```{r}

# Read in the id_to_titles file and add headers
df_id_to_titles <- read.csv("data/id_to_titles.txt", header = FALSE, sep = " ")
colnames(df_id_to_titles) <- c("ID", "Product_Name")

# Sort the in-degree tables from edges$To for the top 10 IDs with the most inbound links
top10_id <- names(head(sort(in_degree, decreasing = TRUE), 10))

# Filter using pipes and print the names of the top 10 products
top10_products <- df_id_to_titles %>% 
  filter(ID %in% top10_id) %>% 
  select(Product_Name) %>% 
  pull()

print(top10_products)


```


